---
title: "Extra credit"
author: "Chimdi Chikezie"
date: "5/31/2023"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(Sleuth3)
library(ggplot2)
library(Rmisc)
library(graphics)
library(leaps)
```


Problem: Fish Market Data Analysis
A dataset was collected on 7 different common fish species in fish market sales. The goal of this dataset was to find the relationship between the weight (Weight, response variable) of fish (in grams) and the height (Height, explanatory variable) of the fish in cm, the diagonal width (Width, explanatory variable) of the fish in cm, the vertical length (Length1, explanatory variable) of the fish in cm, the diagonal length (Length2, explanatory variable) of the fish in cm, the cross length (Length3, explanatory variable) in cm, and the type of the fish (Species, categorical variable with 7 categories). We perform linear regression analysis with this dataset with the goal of explaining the relationship between the response and the explanatory variables. The dataset is attached to the assignment.
```{r}
load("Fish_Data.RData")
head(Data)
#View(Data)
```

(a) Fit the linear regression model of mean weight against all the explanatory variables, but do not include any interaction term in the model. Report the summary() output of the linear model fit.
Plot the residual plot for fitted vs. residual values based on the fitted regression model. What conclusions can you draw from the plot specifically regarding the linear model assumptions? (look for evidence of violations of the model assumptions, identify unusual observations)

```{r}
lin_mod1 <- lm(Weight ~ Species + Length1 + Length2 + Length3 + Height + Width, data = Data)
summary(lin_mod1)
```
From the lm output, the pvalues: 3.64e-11 (Intercept $\beta_0$), 0.035824 (SpeciesParkki $\beta_1$), 0.000311 (SpeciesSmelt $\beta_5$) and 0.029552 (Length1 $\beta_7$) are less than $\alpha = 0.05$ which means that these parameters are statistically significant.
Therefore, we conclude that the means of SpeciesParkki and SpeciesSmelt are different from the mean of other species and that they these sopecies as well as Length1 have an effect on the mean weight of fishes.

```{r}
plot(lin_mod1)
```
From the plot, there is a violation of linearity as there is a u shape made by the observations.
There is a violation of equal variance as the spread of the residuals increases as the fitted values increase.

(b) Transform the response values using the cube-root transformation. You can use the
following code to do the transformation
Data$Weight_new <- (Data$Weight)^(1/3)
Fit the linear regression model of mean transformed weight against all the explanatory
variables, but do not include any interaction term in the model. Report the summary()
output of the linear model fit.
Plot the residual plot for fitted vs. residual values based on the fitted regression model
in part (d). What conclusions can you draw from this plot and comparing this plot with
the residual plot in part (b)?

```{r}
Data$Weight_new <- (Data$Weight)^(1/3)
lin_mod2 <- lm(Weight_new ~ Species + Length1 + Length2 + Length3 + Height + Width, data = Data)
summary(lin_mod2)
```

From the lm output, the pvalues: 3.64e-11 (Intercept $\beta_0$), 0.035824 (SpeciesParkki $\beta_1$), 0.000311 (SpeciesSmelt $\beta_5$) and 0.029552 (Length1 $\beta_7$) are less than $\alpha = 0.05$ which means that these parameters are statistically significant.
Therefore, we conclude that the means of SpeciesParkki and SpeciesSmelt are different from the mean of other species and that they these sopecies as well as Length1 have an effect on the mean weight of fishes.
```{r}
plot(lin_mod2)
```


From the residual plot, there is no violation of linearity as the observations are randomly dispersed around the horizontal line. 
There does not seem to be a violation of equal variance as there is no increase spread of the residuals as the fitted values increase.

c) Based on the transformed weight in part (b), test the hypothesis that the mean transformed
weight of Smelt fish is same as the mean transformed weight of White Fish, keeping the
values of all the other explanatory variables same.

```{r}
new_species <- relevel(Data$Species, ref="Whitefish")
lin_mod3 <- lm(Weight_new ~ new_species + Length1 + Length2 + Length3 + Height + Width, data = Data)
summary(lin_mod3)
```
$H_0:$ mean transformed weight of Smelt fish = mean transformed weight of White Fish
$H_A:$ mean transformed weight of Smelt fish $\neq$ mean transformed weight of White Fish

Since the p-value (9.78e-05) associated with $\beta_6$ (the parameter representing the difference in mean transformed weight of Smelt fish and White Fish) is less than $\alpha = 0.05$. It means $\beta_6$ is significant. 
Therefore, we reject the null hypothesis and conclude that there hat the mean transformed
weight of Smelt fish is not the same as the mean transformed weight of White Fish.

(d) Fit the linear regression model of mean transformed weight against all the explanatory
variables as well as the interaction terms between Species and Height and the interaction
terms between Species and Width. Compare the model in part (b) with this model using
ANOVA F-test. Which model will you choose?
```{r}
lin_mod4 <- lm(Weight_new ~ Species + Length1 + Length2 + Length3 + Height + Width + Species:Height + Species:Width, data = Data)
summary(lin_mod4)
```

```{r}
anova(lin_mod2, lin_mod4)
```

Based on the anova result, the p-value (0.5189) for the f-test comparison is greater than $\alpha = 0.05$ which means it is not significant.
Therefore, we conclude that the interaction terms that were added do not have any effect on the mean transformed weight of fishes and we prefer the simpler model.

(e) For the model in part (b), based on the anova() command output, produce the ANOVA
table with just the three rows of Regression, Residual, and Total.
```{r}
anova(lin_mod2)
```


```{r}
vals <- matrix(c(11, 910.14, 82.74, 146, 6.40, 0.04, 157, 916.54, 5.837), ncol = 3, byrow = TRUE)
colnames(vals) <- c("Df", "SS", "MS")
rownames(vals) <- c("Regression", "Residual", "Total")
vals <- as.table(vals)
vals
```

f) Based on the fitted model in part (b), find the predicted transformed weight of a Bream 
fish of height of 16 cm, diagonal width of 5 cm, vertical length of 25 cm, diagonal length
of 30 cm, cross length of 35 cm. Also, find the 95% prediction interval of the transformed
weight of this fish. If you back-transform (that is, take a cubic power), what is the 95%
prediction interval of the weight of this fish?

```{r}
test_value <- data.frame(Species = "Bream", Length1 = 25, Length2 = 30, Length3 = 35, Height = 16, Width = 5)
pred_weight <- predict(lin_mod2, newdata = test_value)
sprintf("predicted transformed weight = %f", pred_weight)
```

```{r}
sprintf("Prediction interval")
pred_int <-predict(lin_mod2, test_value, interval ="prediction", level=0.95 )
pred_int
```

```{r}
sprintf("Transformed Prediction interval")
pred_int^(3)
```

