---
title: "Chimdi Homework 1"
author: "Chimdi Chikezie"
date: "4/7/2023"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(ggplot2)
library(Rmisc)
library(graphics)
library(Sleuth3)
```

```{r}
if(! (packageVersion("ggplot2") >= "2.0.0")) {
 stop("This version of stat_qqline require ggplot2 version 2.0.0 or greater")
}
```

## Code drawing qqplot and qqline together in ggplot2, you do not need to understand this code
```{r}
StatQQLine <- ggproto("StatQQLine", Stat,
 compute_group = function(data, scales, distribution = qnorm, dparams = list()) {
 data <- remove_missing(data, na.rm = TRUE, "sample", name = "stat_qqline") 
 y <- quantile(data$sample, c(0.25, 0.75))
 x <- do.call(distribution, c(list(p = c(0.25, 0.75)), dparams))
 slope <- diff(y)/diff(x)
 int <- y[1L] - slope * x[1L]
 data.frame(slope = slope, intercept = int)
 },
 required_aes = c("sample")
)

stat_qqline <- function(mapping = NULL, data = NULL, geom = "abline",
 position = "identity", na.rm = FALSE, show.legend = NA,
 distribution = qnorm, dparams = list(),
 inherit.aes = TRUE, ...) {
 layer(
 stat = StatQQLine, data = data, mapping = mapping, geom = geom,
 position = position, show.legend = show.legend, inherit.aes = inherit.aes,
 params = list(na.rm = na.rm, distribution = distribution, dparams = dparams, ...)
 )
}
```

Question 1
Problem 1 (Problem 9.22 in textbook)
Mammal Lifespans and Kleiber’s Law. Kleiber’s law states that the metabolic rate of an
animal species, on average, is proportional to its mass raised to the power of 3/4. The Exercise
8.26 data set contains the mass (in kilograms), average basal metabolic rate (in kilojoules per
day), and lifespan (in years) for 95 mammal species. The data is from A. T. Atanasov, “The
Linear Allo-metric Relationship Between Total Metabolic Energy per Life Span and Body Mass
of Mammals,” Biosystems 90 (2007): 224-33.

(a) 
Draw scatterplots of metabolism versus mass, lifespan versus mass, and lifespan
versus metabolism. Alternatively, you can draw a scatterplot matrix too

# Loading Data
```{r}
#?ex0826
HW1Data <- ex0826
#View(HW1Data)
```

# Plotting the Scatterplot metabolism versus mass
```{r}
qplot(Mass, Metab, data = HW1Data) +  ggtitle('Metabolism versus Mass') + 
  theme(text = element_text(size=12), 
        axis.text.x = element_text(angle=90, vjust=1))
```
# Plot the Scatterplot lifespan versus mass
```{r}
qplot(Mass, Life, data = HW1Data) +  ggtitle('Lifespan versus Mass') + 
  theme(text = element_text(size=12), 
        axis.text.x = element_text(angle=90, vjust=1))
#+ geom_smooth(method='lm', se = FALSE) 
```
# Plot the Scatterplot lifespan versus metabolism
```{r}
qplot(Metab, Life, data = HW1Data) +  ggtitle('Lifespan versus Metabolism') + 
  theme(text = element_text(size=12), 
        axis.text.x = element_text(angle=90, vjust=1))
```

(b) 
Obtain the least squares fit to the linear regression of metabolism on mass,
metabolism on mass to the power of 3/4, and lifespan on mass, separately as simple linear
regression.

```{r}
sprintf("Least squares fit to the linear regression of metabolism on mass")
lin_mod1 = lm(Metab ~ Mass, data = HW1Data)
summary(lin_mod1)
```

```{r}
sprintf("Least squares fit to the linear regression of metabolism on mass to the power of 3/4")
mass <- (HW1Data$Mass)^(3/4)
lin_mod2 = lm(Metab ~ mass, data = HW1Data)
summary(lin_mod2)
```

```{r}
sprintf("Least squares fit to the linear regression of lifespan on mass")
lin_mod3 = lm(Life ~ Mass, data = HW1Data)
summary(lin_mod3)
```

(c) (3 points) Plot the residuals versus the fitted values for each regression. Is there evidence
that the variance of the residuals varies with the fitted values or that there are any outliers?

# Adding the lm results to the data set
```{r}
sprintf("Least squares fit to the linear regression of metabolism on mass")
lin_mod1Data <- fortify(lin_mod1, HW1Data)
lin_mod2Data <- fortify(lin_mod2, HW1Data)
lin_mod3Data <- fortify(lin_mod3, HW1Data)
#View(lin_mod1Data)

# Fitted vs Residual Plot
plot1 <- qplot(.fitted, .resid, data = lin_mod1Data) + geom_hline(aes(yintercept = 0)) + ggtitle('Metabolism on Mass')
plot2 <- qplot(.fitted, .resid, data = lin_mod2Data) + geom_hline(aes(yintercept = 0)) + ggtitle('Metabolism on Mass to the power of 3/4')
plot3 <- qplot(.fitted, .resid, data = lin_mod3Data) + geom_hline(aes(yintercept = 0)) + ggtitle('Lifespan on Mass')

multiplot(plot1, plot2, plot3, cols=2)
```
- There is evidence that there are outliers.For Metabolism on Mass and Metabolism on Mass to the power of 3/4, we can see one value on each plot around 150000 on x axis. While we see a value above 60 on x axis in the Lifespan on Mass plot.

(d) (3 points) Report a summary of each least squares fit in one sentence each, using the
t-test p-values and R2 values.

-Answer

-Metabolism on Mass

The t-value of 65.586 is a statistic that measures the significance of the estimated slope coefficient (Mass) and has a p-value of 2e-16. Since this p-value is less that $\alpha = 0.05$, we reject the null hypothesis that the slope is zero.

The F-statistic = 4302 is measuring the overall significance of the regression model. Since the p value(2.2e-16) associated to it is less that $\alpha = 0.05$, we reject the null hypothesis that there is no relationship between Metabolism and Mass.

The R2 = 0.978 value shows that the model fits the data well and there is a strong relationship between Metabolism and Mass.


-Metabolism on Mass to the power of 3/4
The t-value of 91.895 is a statistic that measures the significance of the estimated slope coefficient (Mass to the power of 3/4) and has a p-value of 2e-16. Since this p-value is less that $\alpha = 0.05$, we reject the null hypothesis that the slope is zero.

The F-statistic = 8445 is measuring the overall significance of the regression model. Since the p value(2.2e-16) associated to it is less that $\alpha = 0.05$, we reject the null hypothesis that there is no relationship between Metabolism and Mass.

The R2 = 0.989 value shows that the model fits the data well and there is a strong relationship between Metabolism and Mass to the power of 3/4.

-Lifespan on mass
The t-value of 5.713 is a statistic that measures the significance of the estimated slope coefficient (Mass) and has a p-value of 1.32e-07. Since this p-value is less that $\alpha = 0.05$, we reject the null hypothesis that the slope is zero.

The F-statistic = 32.64 is measuring the overall significance of the regression model. Since the p value(1.323-07) associated to it is less that $\alpha = 0.05$, we reject the null hypothesis that there is no relationship between Lifespan and Mass.

The R2 = 0.256 value shows that the model does not fit the data well and there is a weak relationship between Lifespan and Mass.



#Question 2 (Problem 9.19 in textbook)

Depression and Education. Has homework got you depressed? It could be worse. Depression,
like other illnesses, is more prevalent among adults with less education than you have.
R. A. Miech and M. J. Shanahan investigated the association of depression with age and
education, based on a 1990 nationwide (U.S.) telephone survey of 2,031 adults aged 18 to
90. Of particular interest was their finding that the association of depression with education
strengthens with increasing age - a phenomenon they called the “divergence hypothesis.”
They constructed a depression score from responses to several related questions. Education
was categorized as (i) college degree, (ii) high school degree plus some college, or (iii) high
school degree only. (See “Socioeconomic Status and Depression over the Life Course,” Journal
of Health and Social Behaviour 41(2) (June, 2000): 162-74.)
Note: This question does not involve any data analysis.


(a) (4 points) Construct a multiple linear regression model in which the mean depression
score changes linearly with age in all three education categories, with possibly unequal
slopes and intercepts. Identify the change in difference of mean depression score between
categories (iii) and (i) with one unit change in age.

- Answer
Multiple linear regression model
Where:
(i) College degree = degree_1
(ii) high school degree plus some college = degree_2
(iii) high school degree = degree_3

$\mu\{depression|age,degree\}=\beta_0 + \beta_1(age) + \beta_2(degree_2) + \beta_3(degree_3) + \beta_4(age * degree_2) + \beta_5(age *degree_3)$

(i) mean depression score: $\beta_0 + \beta_1(age)$
(degree_2 and degree_3 are 0 in this case)

(iii)  mean depression score iii
$\beta_0 + \beta_1(age) + \beta_3 + \beta_5(age)$
$\beta_0 + \beta_3 + (\beta_1 + \beta_5)age$

difference of mean depression score between categories (iii) and (i) = $\beta_5$
                              
(b) (4 points) Modify the model to specify that the slopes of the regression lines with age
are equal in categories (i) and (ii) but possibly different in category (iii). Again identify
the change in difference of mean depression score between categories (iii) and (i) with one
unit change in age.

-Answer

Model: $\beta_0 + \beta_1(age) + \beta_2(degree_2) + \beta_3(degree_3) + \beta_4(age *degree_3)$

(i) mean depression score iv :    $\beta_0 + \beta_1(age)$

(iii) mean depression score ii:  $\beta_0 + \beta_1(age) + \beta_3 + \beta_4(age)$
                        =     $\beta_0 + \beta_3 + (\beta_1 + \beta_4)age$
                        
difference of mean depression score between categories (iii) and (i) = $\beta_4$

(c) (4 points) This and other studies found evidence that the mean depression is high in
the late teens, declines toward middle age, and then increases towards old age. Construct
a multiple linear regression model in which these type of association can be captured by
converting the continuous variable age into a categorical variable with three categories
(late teen, middle age, and old age). Include the three education categories in the model
too. Define the model in such a way that the association between mean depression score
and age variable can be different for different education categories.

-Answer

College degree = degree_0 (reference)
high school degree plus some college = degree_1
high school degree = degree_2
late teen = teen (reference)
middle age = middle
old age = old

$\mu\{depression|age,degree\}=\beta_0 + \beta_1(middle) + \beta_2(old) + \beta_3(degree_1) + \beta_4(degree_2) + \beta_5(middle * degree_1) + \beta_6(old * degree_1) + \beta_7(middle *degree_2) + \beta_8(old *degree_2)$

$\gamma$ $\Gamma$
